---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# cincy

<!-- badges: start -->
[![R-CMD-check](https://github.com/geomarker-io/cincy/workflows/R-CMD-check/badge.svg)](https://github.com/geomarker-io/cincy/actions)
<!-- badges: end -->

The `cincy` package contains simple feature geographic polygon files (`sf` objects) that define Cincinnati Neighborhood, Tract, County, and ZIP Code Geographies.

#### Installing

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("geomarker-io/cincy")
```

#### Data

Data objects are `data.frame`s of class `sf` and are named by geography, source, and optionally vintage:

```{r, echo = FALSE, warning = FALSE}
options(knitr.kable.NA = "")
data(package = "cincy")$results[, "Item"] |>
                        tibble::as_tibble() |>
                        dplyr::rename(name = value) |>
                        tidyr::separate(name, into = c("geography", "source", "vintage"), remove = FALSE) |>
                        dplyr::filter(! name == "dep_index") |>
                        dplyr::arrange(rev(name), vintage) |>
                        dplyr::mutate(name = paste0("`", name, "`")) |>
                        knitr::kable(align = c("lccc")) |>
                        suppressWarnings()
```

Use autocomplete functionality at the `R` prompt (e.g., typing `cincy::` and pressing `TAB` twice) to find the needed `sf` object, narrowing first based on geography (`tract`, `zcta`, `neigh`, `county`), then by source, and then optionally by vintage.

Consult the help file for any object for more information on the data within it:

```{r, eval = FALSE}
?cincy::neigh_ccc
```

Or explore the data documentation online at [https://geomarker.io/cincy/reference](https://geomarker.io/cincy/reference)

#### Examples

Data are returned as simple features objects.

```{r}
library(sf)
```

For example, to get Cincinnati neighborhoods, as defined by the most recent version of community council boundaries from CAGIS:

```{r}
cincy::neigh_ccc
```

Or to get Cincinnati ZIP Codes, as defined by the Census Bureau as ZIP Code Tabulation Areas (ZCTAs), from 2000:

```{r}
cincy::zcta_tigris_2000
```

Alternatively, get the 2010 Counties in Cincinnati's 7-county catchment area:

```{r}
cincy::county_7cc_2010
```



